Abstract Binding Trees

### 2015-10-27 paper.tex
Sort packages
Add support for including agda code
[agda] add variable context; symbols are not terms, but vars are
Add definition of substitution monoidal structures & monoid
@freebroccolo Would you check my work? In generalizing this to our
setting, I sort of just guessed what it "had" to be in order to work,
but you will likely have a better perspective on this.

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{ifthen}
\usepackage{mathtools}
\usepackage{proof}
\usepackage{scalerel}
\usepackage{stmaryrd}
\usepackage{supertabular}
\usepackage{tikz-cd}
\usepackage{unicode-math}
\usepackage{xunicode}

\usepackage{setspace,url,hyperref}

@@

@@

\newcommand\Symbols{\mathbb{I}}
\newcommand\Finsets{\mathbb{F}}
\newcommand\Sorts{\mathcal{S}}
\newcommand\Valences{\mathcal{V}}
\newcommand\Arities{\mathcal{A}}
\newcommand\Operators{\mathcal{O}}

\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}

@@

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\langle #1, #2 \rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle
}
\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

@@

@@

Let $\Symbols$ be an infinite set of symbols. Let $\Finsets$ be the category of
finite subsets of $\Symbols$ and their injective maps; then the comma
construction $\Define{\SCtx}{\Comma{\Finsets}{\Discrete{\Sorts}}}$, with
$\Discrete{\Sorts}$ the discrete category on the set $\mathcal{S}$, is the
category of contexts of symbols, whose objects are finite sets of symbols $U$
and sort-assignments $\Of{\mathfrak{s}}{U\to\Sorts}$, and whose morphisms are
sort-preserving renamings; we will write $\Upsilon$ for a symbol context
$(U,\mathfrak{s})$.

@@

@@

@@

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{agda} % this should come after ams*
\usepackage{catchfilebetweentags}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{ifthen}
\usepackage{mathtools}
\usepackage{proof}
\usepackage{scalerel}
\usepackage{setspace}
\usepackage{stmaryrd}
\usepackage{supertabular}
\usepackage{tikz-cd}
\usepackage{unicode-math}
\usepackage{url}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{xunicode}

@@

% make lualatex play nice with agda
\makeatletter
\ifluatex
   \newwrite\CATCHPATCHOUT
\fi
\def\CatchFBT@Fin@l#1[#2]{%
   \begingroup
      \endlinechar\m@ne \makeatletter #2%
      \ifluatex
         \immediate\openout\CATCHPATCHOUT=\jobname.catchpatch
         \immediate\write\CATCHPATCHOUT {\CatchFBT@tok{\the\CatchFBT@tok}}%
         \immediate\closeout\CATCHPATCHOUT
         \@@input \jobname.catchpatch
      \else
      \scantokens\expandafter{%
         \expandafter\CatchFBT@tok\expandafter{\the\CatchFBT@tok}}%
      \fi
      \CatchFBT@IsAToken{#1}
         {\global#1\expandafter{\the\CatchFBT@tok}}
         {\xdef#1{\the\CatchFBT@tok}}%
      \ifx\CatchFBT@tok#1\else\global\CatchFBT@tok{}\fi
   \endgroup
}% \CatchFBT@Final
\makeatother

@@

\newcommand\Symbols{\mathbf{I}}
\newcommand\SymSets{\mathbb{F}_\Symbols}
\newcommand\FinSets{\mathbb{F}}
\newcommand\Sorts{\mathcal{S}}
\newcommand\Valences{\mathcal{V}}
\newcommand\Arities{\mathcal{A}}
\newcommand\Operators{\mathcal{O}}

\newcommand\Cats{\mathbf{Cat}}
\newcommand\Discrete[1]{#1_{\equiv}}
\newcommand\Sets{\mathbf{Set}}
\newcommand\SCtx{\mathbf{SCtx}}
\newcommand\Ctx{\mathbf{Ctx}}
\newcommand\Seq{\mathbf{Seq}}

\newcommand\OpCat[1]{#1^{op}}

@@

\newcommand\Grothendieck[2]{\oint^{#1} #2#1}
\newcommand\Pair[2]{\langle #1, #2 \rangle}
\newcommand\Tuple[1]{\langle #1 \rangle}
\newcommand\FLift[2]{
  \Operators\langle#1%
  \ifthenelse{\equal{#2}{}}{}{, #2}\rangle
}
\newcommand\PairTimes[2]{\langle #1\, {\scriptstyle {{\times}}}\, #2 \rangle}
\newcommand\SSingleton{\bullet}
\newcommand\Forget[1]{|\, #1\, |}

@@

\DeclareMathOperator\Yoneda{\mathbf{y}}

@@

Let $\Symbols$ be an infinite set of symbols. Let $\SymSets$ be the category of
finite subsets of $\Symbols$ and their injective maps; then the comma
construction $\Define{\SCtx}{\Comma{\SymSets}{\Discrete{\Sorts}}}$, with
$\Discrete{\Sorts}$ the discrete category on the set $\mathcal{S}$, is the
category of contexts of symbols, whose objects are finite sets of symbols $U$
and sort-assignments $\Of{\mathfrak{s}}{U\to\Sorts}$, and whose morphisms are
sort-preserving renamings; we will write $\Upsilon$ for a symbol context
$(U,\mathfrak{s})$.

@@

\begin{definition}
  A \emph{signature} $\Sigma$ is a set of sorts $\Sorts$ together with a copresheaf $\Operators$ of operators.
\end{definition}

@@

\section{Model Theory}

Let $\FinSets$ be the free cocartesian category on a single object (that is,
the category of finite cardinals and functions between them); then the comma
construction $\Define{\Ctx}{\Comma{\FinSets}{\Sorts}}$ is the category of
variable contexts, and $\Define{\Seq}{\SCtx\times\Ctx\times\Sorts}$ is the category of
sequents $\Upsilon\parallel\Gamma\vdash s$.  Then, we fix the presheaf category $\Sets^\Seq$ as
our semantic universe.

\subsection{Signatures and substitution monoidal structures}

Any two presheaves $A,B : \Sets^\Seq$ have a tensor $A\bullet B$,
defined as a coend in the following way:
\[
  \Define{(A\bullet B)(\Upsilon\parallel\Gamma\vdash s)}{
    \int^{\Delta\in\Ctx}
      A(\Upsilon\parallel\Delta\vdash s)
        \otimes \prod_{x\in\Dom\Delta} B(\Upsilon\parallel\Gamma\vdash\Delta(x))
  }
\]

The presheaf of variables $\Define{V(\Upsilon\parallel\Gamma\vdash s)}{\MkSet{x \in \Dom\Gamma
\mid \Gamma(x) = s}}$ is the unit to this tensor, and together they comprise a
substitution monoidal structure. Then, a presheaf $A$ is a monoid if it is
equipped with the following natural transformations:

\[
  \begin{tikzcd}[ampersand replacement = \&]
    {\IMode{V}} \arrow[r, "\IMode\nu"]
    \& {\IMode{A}}
    \& {\IMode{A\bullet A}}
    \arrow[l, "\IMode\varsigma", swap]
  \end{tikzcd}
\]

Intuitively, $\nu$ embeds variables into $A$ and $\varsigma$ equips $A$ with an
operation for simultaneous substitutions of variables.

@@

\section{Code}

\CatchFileBetweenTags\tmp{latex/Code.tex}{agda}

\subsection{Nat}

\ExecuteMetaData[latex/Code.tex]{nat}

\subsection{Fin}

\ExecuteMetaData[latex/Code.tex]{fin}

\subsection{Name}

\ExecuteMetaData[latex/Code.tex]{name}

\subsection{Ctx}

\ExecuteMetaData[latex/Code.tex]{ctx}

\subsection{Signatures}

\ExecuteMetaData[latex/Code.tex]{sign}

\subsection{Trees}

\ExecuteMetaData[latex/Code.tex]{trees}
